Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-03-31 background.tex
finished the rewrite for section 4; made corresponding changes as necessary in section 3

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to a job in a consecutive manner. The assignment first fills up a group, then crosses group boundaries and starts to fill the next group if necessary. As shown in Figure~\ref{fig:dragonfly-overview}, $J2$ gets eight nodes by contiguous placement. Contiguous placement confines the tasks of a job into the same group, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, a job gets a set of nodes randomly selected from all the available nodes in the system. The nodes assigned to $J1$ in Figure~\ref{fig:dragonfly-overview} are attached to different routers in different groups. Routers may be shared by different jobs when random placement is in use. Random placement can distribute the tasks of a job uniformly across the whole system to avoid the possible local congestion. However, random placement may cause possible congestion on global links.

>>>>

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to an application in a consecutive manner. The assignment first fills up a group, then crosses group boundaries and starts to fill the next group if necessary. As shown in Figure~\ref{fig:dragonfly-overview}, $J2$ gets eight nodes by contiguous placement. Contiguous placement confines the tasks of an application into the same group and uses minimum number of routers serving each application, which may result in local network congestion and increase the possibility of hot-spots.

\textbf{Random Placement:} In this policy, a job gets a set of nodes randomly selected from all the available nodes in the system. The nodes assigned to $J1$ in Figure~\ref{fig:dragonfly-overview} are attached to different routers in different groups. Routers may be shared by different jobs and more routers are involved in serving each application when random placement is in use. Random placement can distribute the tasks of an application uniformly across the network to avoid the possible local congestion. However, random placement may cause possible congestion on global links.

### 2016-03-31 study_workload_1.tex
finished the rewrite for section 4; made corresponding changes as necessary in section 3

The study of parallel workload~\Rmnum{1} consists of two parts. Firstly, we analyze the over-all network performance when Workload \Rmnum{1} is running under different placement and routing configurations. Secondly, we isolate each application from the workload, analyze its performance  by scrutinizing the traffic going through the routers that serve each job. The analysis allows us to identify the ``bully" in the workload.

@@

\iffalse

The average communication time spent by all MPI ranks when Workload~\Rmnum{1} is running under different placement and routing configurations shown in Table~\ref{tab:wkld-commtime}. The random placement policy performs comparably with all routing policies and outperforms all contiguous placement configurations. The contiguous placement coupled with minimal routing (CM) results in highest communication time. When coupled with (progressive) adaptive routing (CA, CPA), the communication time can be reduced but still higher than random placement polices.

\fi

@@

%local channel traffic and busy time explanation

%We observe similar phenomena from the local channel traffic and saturated time, shown in Figure \ref{fig:local-channel-traffic}, \ref{fig:local-channel-stime}). Contiguous placement coupled with minimal routing (CM) results in local congestion, and thus the aggregate traffic through local channels of certain routers is quite high. Even with adaptive or progressive adaptive routing, the local congestion caused by grouping all MPI ranks in the same group(s) in contiguous placement can not be eliminated. Random placement is effective at addressing this type of congestion.

@@

@@

As shown in Figure~\ref{fig:mg-commtime}, MultiGrid takes the most communication time when it is running with contiguous placement and minimal routing (CM). When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA) are in use, the communication time is significantly reduced. Random placement coupled with minimal routing (RM) can make further improvement, and reaches the best performance(lowest communication time) when coupled with adaptive and progressive adaptive routing(RA, RPA). Due to the variance of data transfer amount among MPI ranks, MultiGrid communication time also presents great variation, indicated by the long boxes in Figure~\ref{fig:mg-commtime}.

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/lc-traffic}
        \caption{MG Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/lc-traffic}
        \caption{CR Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\\  
    \centering
   \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/gc-traffic}
        \caption{MG Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/gc-traffic}
        \caption{CR Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}
   \caption{The aggregate traffic go through the local and global channels of routers serving specific application. ``CA" and ``CPA" perform comparably, the corresponding lines are overlapped. More routers are involved in serving each application when random placement is in use, compared with contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

\TODO{resume from here}

% Show the evidence from network level, explain why AMG is different. 
We look into network level to identify the culprit behind AMG's abnormal behavior with random placement and adaptive routing. By identifying the compute nodes each MPI rank resides on and routers that serving each application, we can see that the traffic go through the routers serving AMG skyrockets when random placement and adaptive routing is in use, shown in Figure \ref{fig:amg-lc-traffic} and \ref{fig:amg-gc-traffic}. Adaptive routing will redirect the traffic of other applications to go through the routers belonging to AMG, and increase the traffic burden on those routers, thus slow down the communication of AMG.

No router will be shared between applications under contiguous placement. A number of successive groups will be allocated to the same application under contiguous placement policy. When contiguous placement coupled with minimal routing, application will be free of interference from its concurrently running peers. However, application communication traffic will be confined within their allocated groups, thus congestion is very likely to happen depending on application communication intensity. As the lest communication-intensive application in Workload \Rmnum{1}, AMG can benefit from contiguous placement. Under contiguous placement, the traffic go through the local and global channels of the routers belonging to AMG is significantly lower than with random placement, shown in Figure \ref{fig:amg-lc-traffic}, \ref{fig:amg-gc-traffic}.

MultiGrid and CrystalRouter are more communication-intensive compared with AMG. They will suffer congestion on both local and global channels when contiguous placement is in use, shown as the dashed line in Figure \ref{fig:mg-lc-traffic}, \ref{fig:mg-gc-traffic} and Figure \ref{fig:cr-lc-traffic}, \ref{fig:cr-gc-traffic}. On the other hand, the random placement can uniformly distribute their traffic all over the network, alleviate the possible congestion, shown as the solid lines in Figure \ref{fig:mg-lc-traffic}, \ref{fig:mg-gc-traffic} and Figure \ref{fig:cr-lc-traffic}, \ref{fig:cr-gc-traffic}.

\subsection{Key Observations}
In summary, based on the results presented in section \ref{sec: workload-1 network analysis} and \ref{sec: workload-1 app analysis}, we have made the following observations.

@@

\emph{The performance of less communication-intensive jobs in workload will suffer degradation when network random placement and adaptive routing are in use.} AMG in Workload \Rmnum{1} are ``bullied" by their concurrently running communication-intensive peers, MultiGrid and CrystalRouter. AMG shares routers and groups with MultiGrid and CrystalRouter when random placement is in use. The traffic from MultiGrid and CrystalRouter will be redirected to the routers that serving AMG, slowing down AMG's communication. \footnote{We have tried three different congestion sensing schemes that used in adaptive routing\cite{won-prog-adaptive}, although there are some variations between the results, none of the congestion sensing scheme can prevent the ``bully" from happening.}

\emph{Consistent performance of each application can be guaranteed when contiguous placement and minimal routing are in use.} The router and group sharing between jobs are prohibited in contiguous placement policy. The traffic won't be redirected from congested routers to less busy ones when minimal routing is in use. Thus the interference between concurrently running jobs will be eliminated and the ``bully" behavior will be terminated. 
%However, contiguous placement and minimal routing usually results in local congestion and unbalanced network utilization.

>>>>

The study of parallel workload~\Rmnum{1} consists of two parts. Firstly, we analyze the over-all network performance when Workload~\Rmnum{1} is running under different placement and routing configurations. Secondly, we isolate each application from the workload, analyze its performance by scrutinizing the traffic going through the routers that serve each job. The analysis allows us to identify the ``bully" in the workload.

@@

We study the network performance under different placement and routing configurations by analyzing the traffic going through the routers and their saturated time. 
The distribution of traffic over the network can reveal the loading situation, 
indicating whether the network is load-balanced or not. 
The saturated time can indicate the congestion situation on certain part of the network.

@@

@@

%===========================================================
%from the workload perspective to study the network performance. 
The efficiency of the workload communication is another perspective to evaluate the network performance. 
The average communication time spent by all MPI ranks when Workload~\Rmnum{1} is running under different placement and routing configurations shown in Table~\ref{tab:wkld-commtime}. 
The random placement policy performs comparably with all routing policies (RM, RA and RPA). They outperform all contiguous placement configurations (CM, CA and CPA). The contiguous placement coupled with minimal routing (CM) results in highest communication time. When coupled with (progressive) adaptive routing (CA, CPA), the average communication time can be greatly reduced. When random placement is in use, the workload communication efficiency can be further improved.

The random placement can balance the network traffic load by uniformly distribute the traffic. The (progressive) adaptive routing can avoid the hot-spots by redirecting the traffic away from congested area. The network can reach its optimal performance, when random placement coupled with (progressive) adaptive routing are in use.

@@

As shown in Figure~\ref{fig:mg-commtime}, 
MultiGrid takes the most communication time when it is running with contiguous placement and minimal routing (CM). 
When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA) are in use, 
the communication time is significantly reduced. 
Random placement coupled with minimal routing (RM) can make further improvement, 
and reaches the best performance(lowest communication time) when coupled with adaptive and progressive adaptive routing(RA, RPA). 
Due to the variance of data transfer amount among MPI ranks, 
MultiGrid communication time also presents great variation, 
indicating by the long boxes in Figure~\ref{fig:mg-commtime}.

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/lc-traffic}
        \caption{MultiGrid Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/lc-traffic}
        \caption{CrystalRouter Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\\  
    \centering
   \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/gc-traffic}
        \caption{MultiGrid Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/gc-traffic}
        \caption{CrystalRouter Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}
   \caption{The aggregate traffic go through the local and global channels of routers serving specific application. ``CA" and ``CPA" perform comparably, the corresponding lines are overlapped. More routers are involved in serving each application when random placement is in use, compared with contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

% Show the evidence from network level, explain why AMG is different. 
We look into network level to identify the culprit behind AMG's abnormal behavior with random placement. 
By identifying the compute nodes that each MPI rank resides on and the routers that serving each application, 
we analyze the traffic going through the routers belonging to each application and 
present the results in Figure~\ref{fig:3app-lc-gc-traffic}. 
Compared with contiguous placement, 
the number of routers serving each application when random placement is in use are greatly increased, 
thus the solid lines are longer than dashed lines as shown in Figure~\ref{fig:3app-lc-gc-traffic}.

The routers serving MultiGrid and CrystalRouter have high volume of traffic on both local and global channels when contiguous placement is in use, 
shown as the dashed line in Figure~\ref{fig:mg-lc-traffic},~\ref{fig:mg-gc-traffic} and Figure~\ref{fig:cr-lc-traffic},~\ref{fig:cr-gc-traffic}. 
On the other hand, the random placement can uniformly distribute the traffic of MultiGrid and CrystalRouter over the network, 
alleviate local congestion and balance the traffic load among all routers, 
shown as the solid lines in Figure~\ref{fig:mg-lc-traffic},~\ref{fig:mg-gc-traffic} and Figure~\ref{fig:cr-lc-traffic},~\ref{fig:cr-gc-traffic}. 
MultiGrid and CrystalRouter benefit from random placement by getting their high volume of traffic being redirected 
from their over-loaded routers to the less busy routers in the network. 
In this case, a majority of those less busy routers are belonging to AMG.

Since AMG is less communication-intensive, 
the routers serving AMG have small volume of traffic on local and global channels when the contiguous placement is use, 
shown as the dashed line in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
When random placement is in use, 
more routers are involved in serving AMG, 
and the traffic skyrockets on the local and global channels, 
shown as the solid lines in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
The increased traffic comes from the communication-intensive MultiGrid and CrystalRouter.

AMG behaves abnormally when random placement is in use 
because the routers serving AMG have to take the traffic from MultiGrid and CrystalRouter. 
The extra traffic from the other communication-intensive applications 
takes advantage the network resources belonging to AMG and slows down AMG's communication. 
We refers this phenomenon as AMG being ``bullied" by MultiGrid and CrystalRouter.

\subsection{Key Observations}
In summary, based on the results presented in section~\ref{sec: workload-1 network analysis} and~\ref{sec: workload-1 app analysis}, we have made the following observations.

@@

\emph{The performance of less communication-intensive jobs in workload suffers degradation when random placement and adaptive routing are in use.} 
AMG in Workload~\Rmnum{1} are ``bullied" by its concurrently running communication-intensive peers, MultiGrid and CrystalRouter. 
AMG shares routers and groups with MultiGrid and CrystalRouter when random placement is in use. 
The traffic from MultiGrid and CrystalRouter is redirected to the routers that serving AMG, 
slowing down AMG's communication. 
\footnote{We have tried three different congestion sensing schemes that used in adaptive routing~\cite{won-prog-adaptive}, although there are some variations between the results, none of the congestion sensing scheme can prevent the ``bully" from happening.}

\emph{The consistency performance of each application can be guaranteed when contiguous placement and minimal routing are in use.} 
The router and group sharing among applications are prohibited in contiguous placement policy. Minimal routing will not redirect traffic from congested routers to less busy ones.
Thus the interference between concurrently running jobs will be eliminated and the ``bully" behavior will be terminated.
